Supplementary figures

### 2021-07-01 code/supplemental_v2.Rmd
Add supplemental fig5. Model comparison with other classification. ecdf and error distribution

---
title: "Supplementary figures"
output:
  word_document:
    reference_docx: "../text/word-style-reference-ttmanu-iscb-iscience-v2.docx"
always_allow_html: TRUE
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE, message = FALSE)
suppressPackageStartupMessages(c(
  library(tidyverse),
  library(data.table),
  library(plot3D),
  library(flextable),
  library(knitr),
  library(kableExtra),
  library(xlsx)
))
```

@@

```{r supplemental-table-st1-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")

## Parameters ##
n_path <- 377 ## how many top pathways do you want to see. 1 = top path, 2 = top path and second most
mutated_gene_position <- c(1,2,3) ## what gene do you want to see. 1 = top gene, 2 = second most, 3 = third
show_perc <- TRUE ## show perc for mutated gene

toppath <- gdc_reactome_path_analysis %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:n_path,] %>% .[, "path_name"] ## getting the top mutated pathways

id_to_names <- gene2level2_lookup %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "PATH_ID") %>% ## convert path_id to path_name
  left_join(., pathname2level2, by = "PATH_ID") %>%
  remove_rownames() %>% column_to_rownames(var = "PATH_NAME") %>%
  dplyr::select(-PATH_ID) %>%
  t() %>% as.data.frame()

gene_mutation <- gdc_gene_exp_adj %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id)

table_st1 <- list() ## list for forloop

for(n in toppath) {
  message(n)
  ## finding top mutated gene in pathway
  gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
  top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, ] %>% ## only include genes in paths
      rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>% ## sum the pathway mutation
    rownames_to_column(var = "gene") %>%
      mutate(n_sample = length(gene_mutation),
             perc_mut = round((obs_count / length(gene_mutation)) * 100, digits = 1)) %>% 
       arrange(desc(obs_count))
  top_gene_in_path <- list()
  for(p in mutated_gene_position) {
    message(p)
    if(show_perc == TRUE) {
    top_gene_in_path[[p]] <- paste0(top_mut_gene$gene[p], " (", top_mut_gene$perc_mut[p],"%)")  
    }
    if(show_perc == FALSE) {
    top_gene_in_path[[p]] <- paste0(top_mut_gene$gene[p])  
    }
  }
  top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% ## rbind list
   'rownames<-' (mutated_gene_position) %>% t() %>% as.data.frame()
  ## calculating the percent mutated samples
  spec_path <- data.frame(Pathway = n,
                          Percent_Mutated = round(gdc_reactome_path_analysis %>% .[n,] %>% sum() / length(gdc_reactome_path_analysis) * 100, digits = 1))
  table_st1[[n]] <- cbind(spec_path, top_gene_in_path)
}
table_st1 <- as.data.frame(do.call(rbind, table_st1))
save(table_st1, file = "../data/table_st1.Rda")  
rm(list = ls(all.names = T))  
gc()
```

```{r supplemental-table-st1, eval = FALSE}
load("../data/table_st1.Rda")
## for html only
# table_st1 %>% dplyr::select(-Pathway) %>% knitr::kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
#   column_spec(1, bold = T, color = "black")

## for word doc
# table_st1 %>% flextable() %>% 
#   set_caption(caption = "") %>%
#   align(part = "all") %>% 
#   width(., width = c(4,0.5,0.5,0.5,0.5)) %>% 
#   fontsize(size = 8, part = "body")
```

```{r st2, fig.width= 9, eval = FALSE}
##**Supplementary Table 2. Cluster-specific enriched pathways (effect size >= 0.30).**
## docx friendly
load("../data/moderate_enrich_prob.Rda")

table_st2 <- moderate_enrich_prob %>% filter(difference >= 0.30 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

table_st2$log10pvalue <- round(-log10((table_st2$Signif + 1) / 1E5), digits = 2) ## Getting p value ## adding 1 for pseudo count.. 1E5 is the n simulations performed
table_st2$Enrichment_Score <- round(table_st2$Enrichment_Score, digits = 2)
table_st2$Signif <- NULL ## we don't need this anymore

# st1 <- st1 %>% flextable() %>% 
#   set_caption(caption = "") %>%
#   align(part = "all") %>% 
#   width(., width = c(4,0.75,1,1)) %>% 
#   fontsize(size = 10, part = "body")
#

# names(st1) <- c("Pathway", "Cluster", "Effect Size", "-log(p)")
#  st1 %>%  knitr::kable(format = "latex", longtable = TRUE, align = c("l", rep("c",3))) %>%
#    kable_styling(latex_options=c("scale_down", "repeat_header"), full_width = FALSE, font_size = 12) %>%
#    column_spec(1, color = "black", width = "4in")

## Saving st1 and st2 in excel

write.xlsx(table_st1, file="supplemental.table.xlsx", sheetName="ST1", row.names=FALSE)
write.xlsx(table_st2, file="supplemental.table.xlsx", sheetName="ST2", append=TRUE, row.names=FALSE)
```

>>>>

---
title: "Supplementary figures"
output:
  word_document:
    reference_docx: "../text/word-style-reference-ttmanu-iscb-iscience-v2.docx"
always_allow_html: TRUE
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE, message = FALSE)
suppressPackageStartupMessages(c(
  library(tidyverse),
  library(data.table),
  library(plot3D),
  library(flextable),
  library(knitr),
  library(kableExtra),
  library(xlsx),
  library(patchwork)
))
```

@@

```{r supplemental-s5-plot-data, eval = FALSE}
## Calculate error distribution

## break error into bins for entropy calculation
## Set number of bins - 1024 bins
binsize <- 1024 - 2 ## -2 bins for the less than -3650 and greater than 3650 values
evaluebreak <- c(-1e10,seq(from = -3650.00, to = 3650.00, by = 3650*2/binsize), 1e10)
elabel <- list()
counter <- 2
for(i in evaluebreak[c(-1,-(length(evaluebreak)-1))]) {
  if(i == -1e10) {
    elabel[[counter]] <- paste0(i, "+")
  } else if(i == 1e10) {
    elabel[[counter]] <- paste0(i, "+")
  } else {
    elabel[[counter]] <- paste0(i, "-", round(evaluebreak[[counter+1]], digits = 2))
  }
  counter <- counter + 1
}
elabel <- as.character(elabel)
elabel[1] <- "-3650+"
elabel[1024] <- "3650+"

load("../data/deadpat_all.rda")
deadpat <- deadpat %>% rownames_to_column(var = "stanid")
## convert negative tdeath to 0
deadpat[which(deadpat$tevent < 0), "tevent"] <- 0
model <- c("knn", "icluster", "coca", "tis")
tdeath.error <- list() ## store error values
entropy <- list() 
for(n in model) {
  ## Load correct model
  if(n == "knn") {
    load("./survival/survmodel_effcl_knn_062821.stansave")
  } else if(n == "icluster") {
    load("./survival/survmodel_effcl_icluster_062821.stansave")
  } else if (n == "coca") {
    load("./survival/survmodel_effcl_coca_062821.stansave")
  } else if (n == "tis") {
    load("./survival/survmodel_effcl_tis_062821.stansave")
  } else {
    print("Model does not exist")
  }
  ## Extract tdeath for "adeath" indivduals
  tdeath <- samps %>% as.data.frame(.) %>% .[, grep(pattern = "tdeath", x = names(.))]
  names(tdeath) <- gsub("tdeath\\[(.*)\\]","\\1", names(tdeath)) ## update colname with stanid
  tdeath <- tdeath %>% gather(data = ., key = "stanid", value = "predicted") %>%
    left_join(x = ., y = deadpat[,c("stanid", "tevent", "eventtype")], by = "stanid") %>%
    dplyr::filter(eventtype == "adeath")
  ## Calculate error for each sample
  tdeath$error <- tdeath$predicted - tdeath$tevent
  tdeath.error[[n]] <- tdeath$error
  ## Split error data into bins and tally up 
  tdeath <- tdeath %>% mutate(ebin = cut(x = error,
                                         breaks = evaluebreak,
                                         right = T,
                                         labels = elabel))
  eprob <- as.data.frame(table(tdeath$ebin))
  eprob <- eprob %>% dplyr::filter(Freq > 0) %>%
    mutate(prob = Freq / nrow(tdeath))
  ## Entropy calculation
  entropy[[n]] <- round(-sum(eprob$prob*log(eprob$prob))/log(2), digit = 3)
}
knn <- data.frame(error = tdeath.error$knn,
                  model = "knn")
icluster <- data.frame(error = tdeath.error$icluster,
                       model = "icluster")
coca <- data.frame(error = tdeath.error$coca,
                   model = "coca")
tis <- data.frame(error = tdeath.error$tis,
                   model = "tis")
tdeath.error <- rbind(knn, icluster, coca, tis); rm(knn, icluster, coca, tis); gc()

## empirical cumulative distribution function
## Model selection
model <- c("knn", "icluster", "coca", "tis")

for(n in model) {
   ## Load correct model
  if(n == "knn") {
    load("./survival/survmodel_effcl_knn_062821.stansave")
  } else if(n == "icluster") {
    load("./survival/survmodel_effcl_icluster_062821.stansave")
  } else if (n == "coca") {
    load("./survival/survmodel_effcl_coca_062821.stansave")
  } else if (n == "tis") {
    load("./survival/survmodel_effcl_tis_062821.stansave")
  }
  else {
    print("Model does not exist")
  }
  dfs <- as.data.frame(samps) %>% 
    select(starts_with("tdeath[")) %>%
    stack(.) %>%
    'colnames<-' (c("tdeath", "stanid")) %>%
    .[sample(x = nrow(.), size = 1e5),]
  assign(x = n, value = dfs)
}

save(entropy, tdeath.error, knn, icluster, coca, tis, file = "../data/s5_plot.rda")
```

```{r supplemental-s5-plot, fig.width=9}
load("../data/s5_plot.rda")
load("../data/deadpat_all.rda")

s5.1 <- ggplot(knn) + stat_ecdf(aes(x=tdeath),color="#00BFC4", size = 1) + 
  stat_ecdf(aes(x=tevent),data=subset(deadpat,eventtype=="adeath"),col="#00BA38", size = 1) +
  stat_ecdf(aes(x=tdeath),data=icluster, color="#7CAE00", size = 1) + 
  stat_ecdf(aes(x=tdeath),data=coca, color="#F8766D", size = 1) + 
  stat_ecdf(aes(x=tdeath),data=tis, color="#F564E3", size = 1) + 
  coord_cartesian(xlim=c(0,365*5)) + 
  labs(x = "Survival Time (years past diagnosis)", y = "Fn(Survival Time)") +
  annotate(geom = "text", x = 250, y = 0.5, label = "Actual Death", color = "#00BA38", size = 3) +
  theme_minimal()

s5.2 <- ggplot(data = tdeath.error, aes(x = error, group = model, fill = model)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  scale_x_continuous(breaks=seq(-4000, 4000, by = 2000), limits = c(-3650,3650)) +
  labs(x = "Error (Predicted - Actual)", y = "Density")#, 
       #subtitle = paste0("Entropy - knn: ", entropy$knn, " ; icluster: ", entropy$icluster, "\ncoca: ", entropy$coca, " ; tis: ", entropy$tis))

layout = "
AABB
AABB
"

s5.1 + s5.2 + plot_annotation(tag_levels = "A") + plot_layout(design = layout)
```

**S5 Fig. Performance comparison of different classification model - knn, COCA, iCluster, tissue-of-origin (tis).** A) Empirical cumulative distribution function comparing time of death prediction of all classification models with actual time of death (dark green). B) Error distribution plot of predicted and actual time of death.
